---
title: "Clustering and heatmap"
author: "R. Stam"
date: "19 November 2014"
output: html_document
---
Loading required libraries and data
------------------------------
```{r,warning=FALSE}
library(gplots) 
library(ggplot2)
library(RColorBrewer)
library(reshape)
DEG2 <- as.matrix(read.table("EdgeR_DEG2.txt"))

```

Hierarchical clustering and draw heatmap
------------------------------
hclust to do clustering. Dist. matrix is correlation.
Also, all rows with no values are removed (just in case).
The tree is cut at tree lenghth / 1.5. This generates 10 clusters.
```{r,warning=FALSE}
#DEG2 <- DEG2[rowMin(as.matrix(DEG2))>0,]
hr <- hclust(as.dist(1-cor(t(DEG2), method="pearson")), method="complete") # Cluster rows by Pearson correlation.
mycl <- cutree(hr, h=max(hr$height)/1.5) # Cut the tree at specific height and color the corresponding clusters in the heatmap color bar.
mycolhc <- brewer.pal(9,"RdYlBu")
mycolhc <- mycolhc[as.vector(mycl)]
heatmap(DEG2, Rowv=as.dendrogram(hr), Colv=NA, scale="row", col=colorRampPalette(brewer.pal(9,"Blues"))(100), RowSideColors=mycolhc) 

```


Plotting expression profiles for all clusters
--------------------------
Data are extracted from clustering objects.
Expression value files are specified.


```{r}
cut <- as.data.frame(mycl)
cut$gene <- rownames(cut)
colnames(cut) <- (c("cluster","gene"))
data <- DEG2
```

Create a function to plot as many linegraphs as needed with all genes for a certain cluster in one graph

```{r,warning=FALSE}
#PLOT CLUSTERS
multiplot <- function(..., plotlist=NULL, cols) {
  require(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # Make the panel
  plotCols = cols                          # Number of columns of plots
  plotRows = ceiling(numPlots/plotCols) # Number of rows needed, calculated from # of cols
  
  # Set up the page
  grid.newpage()
  pushViewport(viewport(layout = grid.layout(plotRows, plotCols)))
  vplayout <- function(x, y)
    viewport(layout.pos.row = x, layout.pos.col = y)
  
  # Make each plot, in the correct location
  for (i in 1:numPlots) {
    curRow = ceiling(i/plotCols)
    curCol = (i-1) %% plotCols + 1
    print(plots[[i]], vp = vplayout(curRow, curCol ))
  }
  
}

```

Set all parameters required and loop over all things to draw plots and add to allPlots object
```{r,warning=FALSE}

clusters <- c(1:10)
colln <- 5
breaks <- axTicks(side=1)
times <- c(0,4,8,16,24)
allPlots <- list()


for (i in c(seq(1,length(clusters), by=1))){
  # extract the genes for the cluster
  cluster <- cut[cut$cluster == clusters[i],]
  clustergenes <- as.vector(cluster$gene)
  cluslines <- data[clustergenes,]
  
  # melt the df
  colnames(cluslines) <-times
  df.melted <- melt(t(cluslines))
  colnames(df.melted) <- c("times", "gene", "value")
  df.melted$times <- factor(df.melted$times, levels=unique(df.melted$times))
  # make plot variable
  pl <- ggplot(data = df.melted, aes(x = times, y = value, colour = gene, group = gene)) +
    geom_line() +
    scale_y_continuous(breaks=breaks)
  allPlots[[i]] <- pl + theme(legend.position = "none")
         
}

multiplot(plotlist=allPlots, cols = colln)
```

Alternatively, plot boxplots which represent expression value spread for each of the clusters
```{r, warning=FALSE}
DGE <- DEG2
colnames(DGE) <- c(0,4,8,16,24)
ct2 <- cutree(hr, h=max(hr$height)/1.5)
augDGE <- merge(DGE,as.data.frame(ct2),by="row.names")
meltDGE <- melt(augDGE,c("Row.names","ct2"))
ggplot(meltDGE,aes(x=as.factor(variable),y=log2(value))) + facet_grid(.~ct2) + geom_boxplot()
```

